Using reference models in variable selection \vspace{.1in}

### 2019-12-22 paper/ref_approach.tex
minor edits

    \author{
    Federico Pavone \footnote{Department of Computer Science, Aalto University, Finland.} $\,$\footnote{Correspondence concerning this article should be addressed to TODO.}, 
  Juho Piironen$^*$,
  Paul-Christian B\"{u}rkner$^*$
  and Aki Vehtari$^*$
  }

@@

\cite{paper:bodyfat} report results using steplm with a significance
level of 0.157 with AIC selection \citep{akaike1974new}, fixing
abdomen and height to be always included in the model. For better
comparison we don't fix any of the variables.  The steplm approach is
carried out combining the \texttt{step} and \texttt{lm} functions in R.

@@

The inclusion frequencies of each variable in the final model given 100
bootstrap samples are shown in Figure
\ref{fig:inclusion_frequencies}. In case of projpred there are two
variables, `abdomen' and `weight', which have inclusion frequencies
above 50\% (`abdomen' is the only one included always), the third most
frequently included is `wrist' at 44\%, and the fourth one is `height'
at 35\%.  The steplm approach has seven variables with inclusion
frequencies above 50\%. Such a higher variability and lower stability
of steplm can be observed also in the bootstrap model selection
frequencies reported in Table \ref{tab:model_frequencies}. For
example, the first five selected models have a cumulative frequency of
76\% with projpred, but only of 14\% with steplm. In addition, the
sizes of the selected models with projpred are much smaller with
steplm.
\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/inc_prob.pdf}
  \caption{Bootstrap inclusion frequencies after 100 bootstrap samples.}
  \label{fig:inclusion_frequencies}
\end{figure}

@@

Figure \ref{fig:bodyfat_sensitivity_vs_fdr} shows the sensitivity
against the false discovery rate.  In almost all of the bootstrapped
subsamples, the reference
model improves the selection both in terms of sensitivity and false
discovery rate. When $n=50$, we observe worse false
discovery rates, yet by a lower amount compared to the gain in
sensitivity. Again, we observe that the benefits are more evident as
the selection becomes more challenging (i.e., lower number of
observations).
The great performance of projpred in minimal subset selection is not
carried over for the complete variable selection with iterative
projpred (even when changing the reference model into an full encompassing 
linear regression with regularised horseshoe prior), and the methods 
specifically designed for the complete variable selection perform better. 
Figure \ref{fig:bodyfat_stability} shows the stability
results using the measure by
\cite{paper:stability}. The benefits of the reference model are here
marginal with only small improvements.

@@

The projective prediction approach was not designed for the complete
variable selection. We tested a simple iterative version of projpred,
with mixed results and the methods specifically designed for complete
variable selection (especially loc.fdr) performed better in our
experiments. It is left for future resarch to develop 
a better projective prediction approach for complete variable
selection problems.

>>>>

    \author{
    Federico Pavone \footnote{Department of Computer Science, Aalto University, Finland.} $\,$\footnote{Correspondence concerning this article should be addressed to Federico Pavone, Department of Decision Sciences, Bocconi University, Via R\"{o}ntgen 1, 20136 Milano, Italy. E-mail: federicopavone01@gmail.com.}, 
  Juho Piironen$^*$,
  Paul-Christian B\"{u}rkner$^*$
  and Aki Vehtari$^*$
  }

@@

\cite{paper:bodyfat} report results using steplm with a significance
level of 0.157 with AIC selection \citep{akaike1974new}, fixing
abdomen and height to be always included in the model. For better
comparison we do not fix any of the variables.  The steplm approach is
carried out combining the \texttt{step} and \texttt{lm} functions in R.

@@

The inclusion frequencies of each variable in the final model given 100
bootstrap samples are shown in Figure
\ref{fig:inclusion_frequencies}. In case of projpred there are two
variables, `abdomen' and `weight', which have inclusion frequencies
above 50\% (`abdomen' is the only one included always), the third most
frequently included is `wrist' at 44\%, and the fourth one is `height'
at 35\%.  The steplm approach has seven variables with inclusion
frequencies above 50\%. Such a higher variability and lower stability
of steplm can be observed also in the bootstrap model selection
frequencies reported in Table \ref{tab:model_frequencies}. For
example, the first five selected models have a cumulative frequency of
76\% with projpred, but only of 14\% with steplm. In addition, the
sizes of the selected models with projpred are much smaller than the ones 
selected with steplm.
\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/inc_prob.pdf}
  \caption{Bootstrap inclusion frequencies after 100 bootstrap samples.}
  \label{fig:inclusion_frequencies}
\end{figure}

@@

Figure \ref{fig:bodyfat_sensitivity_vs_fdr} shows the sensitivity
against the false discovery rate.  In almost all of the bootstrapped
subsamples, the reference
model improves the selection both in terms of sensitivity and false
discovery rate. When $n=50$, we observe worse false
discovery rates, yet by a lower amount compared to the gain in
sensitivity. Again, we observe that the benefits are more evident as
the selection becomes more challenging (i.e., lower number of
observations).
The great performance of projpred in minimal subset selection is not
carried over for the complete variable selection with iterative
projpred (even changing the reference model with a full encompassing 
linear regression with regularised horseshoe prior), and the methods 
specifically designed for the complete variable selection perform better. 
Figure \ref{fig:bodyfat_stability} shows the stability
results using the measure by
\cite{paper:stability}. The benefits of the reference model are here
marginal with only small improvements.

@@

The projective prediction approach was not designed for the complete
variable selection. We tested a simple iterative version of projpred,
with mixed results and the methods specifically designed for complete
variable selection (especially loc.fdr) performed better in our
experiments. It is left for future research to develop 
a better projective prediction approach for complete variable
selection problems.
